### 2019-03-16 Chapters/Chapter1.tex
minor thesis changes

Our goal in this project is to investigate the techniques to manipulate the magnetic state in an antiferromagnetic insulator, which could potantially lead to applications for ultra-fast writing devices. For ferromagnetic materials it has already been shown that short laser pulses can induce an effective magnetic field that can reach an amplitude of a few Tesla (\cite{Qaiumzadeh2016}, \cite{Qaiumzadeh2013}).

>>>>

Our goal in this project is to investigate the techniques to manipulate the magnetic state in an antiferromagnetic insulator, which could potantially lead to applications for ultra-fast writing devices. For ferromagnetic materials it has already been shown that short laser pulses can induce an effective magnetic field that can reach an amplitude of a few Tesla \cite{Qaiumzadeh2016}, \cite{Qaiumzadeh2013}.

### 2019-03-16 Chapters/Chapter2.tex
Add more references and short summary in the intro
minor thesis changes

\chapter{Basic models for solid state}

@@

Where $\hat{c}^{\dagger}_{i\sigma}$ ($\hat{c}_{i\sigma}$) creates (annihilates) an electron in site $i$ and spin state $\sigma$. The quantity $t$ is the hopping amplitude, and plays the same role $B$ in the first quantization formalism. We can obtain back \ref{TBBand} by changing basis to momentum space $\hat{c}_{k\sigma} = \frac{1}{\sqrt{N}}\sum_j e^{-i\bs{k}\cdot\bs{r}_j}\hat{c}_{j\sigma}$.

@@

As we discuss earlier, even tough the tight binding model is successful in describing the band structure of most solids, it fails in some cases. In strongly correlated electron systems, the effect of electron-electron Coulomb interactions is important, and this effect is neglected in \ref{TBModel}. The simplest way to take this effect into account is to consider the interaction of electrons in the same site only, that is, adding a term of the form $\text{U}\sum_i \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$:
\begin{equation}
\label{HH}
\hat{H} = -t\sum_{\langle i,j \rangle \sigma} \hat{c}^{\dagger}_{i\sigma}\hat{c}_{j\sigma} + \text{U}\sum_i \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation}
This new term will add an energy $+\text{U}$ when a site is double occupied ($\hat{n}_{i\uparrow} = \hat{n}_{i\downarrow} = 1$) and will not contribute in the more normal case, that $\hat{n}_{i\uparrow} + \hat{n}_{i\downarrow} = 1$. The half-filled case is most commonly studied, this is the situation where the system has one electron per site (or it has half the maximum allowed number). This situation is simpler in many ways and it exhibits a lot of interesting phenomena (Mott insulating behavior, antiferromagnetic order, etc.). However, other cases have also been widely studied.

Notice that if we set $\text{U}=0$ in \ref{HH} we recover the tight binding Hamiltonian, therefore we can say that the Hubbard model is an expansion of the tight binding model. The opposite limit $t=0$ is more interesting. In the half-filling case, for $\text{U}>0$ the ground state will be any state with exactly one electron per site. Since each electron can be in two spin states there is a $2^N$ degeneracy.

>>>>

\chapter{Basic models for solid state}
\label{ChapBasicModels}

@@

Where $\hat{c}^{\dagger}_{i\sigma}$ ($\hat{c}_{i\sigma}$) creates (annihilates) an electron in site $i$ and spin state $\sigma$. The quantity $t$ is the hopping amplitude, and plays the same role $B$ in the first quantization formalism. We can obtain back \ref{TBBand} by changing basis to momentum space $\hat{c}_{k\sigma} = \frac{1}{\sqrt{N}}\sum_j e^{-i\bs{k}\cdot\bs{r}_j}\hat{c}_{j\sigma}$. In Appendix \ref{APD} we use this formalism to compute the band structure of a two dimensional system in a honeycomb lattice (i.e. graphene).

@@

As we discuss earlier, even tough the tight binding model is successful in describing the band structure of most solids, it fails in some cases. In strongly correlated electron systems, the effect of electron-electron Coulomb interactions is important, and this effect is neglected in \ref{TBModel}. The simplest way to take this effect into account is to consider the interaction of electrons in the same site only, that is, adding a term of the form $\text{U}\sum_i \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$:
\begin{equation}
\label{HH}
\hat{H} = -t\sum_{\langle i,j \rangle \sigma} \hat{c}^{\dagger}_{i\sigma}\hat{c}_{j\sigma} + \text{U}\sum_i \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation}

This is known as the Hubbard model, which was first proposed by Hubbard \cite{Hubbard1963} to describe electron correlations in $d$ or $f$ bands in transition metals. The result was a simplified model to predict metal-insulator transitions. However, it exhibits much richer phenomena. It can be considered an impoved tight binding model, describing Mott insulators, which cannot be describied by band theory (i.e. they should be conductors according to band theory, but electron-electron interactions keep away from). Also, the attractive Hubbard model (where $\text{U} < 0$) can be considered an effective model relevant for understanding high temperature superconductivity \cite{Alexandrov1981}. Despite its simplicity the ground state has no analytic form in more than one dimensions, and different computational methods must be used to study it.

 Figure \ref{fig:Hubbard} shows a graphical representation of the parameters in \ref{HH}. The new term $\text{U}\sum_i \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$ will add an energy $+\text{U}$ when a site is double occupied ($\hat{n}_{i\uparrow} = \hat{n}_{i\downarrow} = 1$) and will not contribute in the more normal case, that $\hat{n}_{i\uparrow} + \hat{n}_{i\downarrow} = 1$. The half-filled case is most commonly studied, this is the situation where the system has one electron per site (or it has half the maximum allowed number). This situation is simpler in many ways and it exhibits a lot of interesting phenomena (Mott insulating behavior, antiferromagnetic order, etc.). However, other cases have also been widely studied.

\begin{figure}
\centering
  \includegraphics[width=\linewidth]{../Figures/hubbard.png}
  \caption{Image from Scalettar's notes representing the kinetic term $t$ and the on-site interaction $\text{U}$ in the Hubbard model.} 
\label{fig:Hubbard}
\end{figure}

Notice that if we set $\text{U}=0$ in \ref{HH} we recover the tight binding Hamiltonian, therefore we can say that the Hubbard model is an expansion of the tight binding model. The opposite limit $t=0$ is more interesting. In the half-filling case, for $\text{U}>0$ the ground state will be any state with exactly one electron per site. Since each electron can be in two spin states there is a $2^N$ degeneracy. The subspace spanned by these states is referred to as low energy subspace, and in this case is has energy $E_0 = 0$. If we switch $0 \neq t \ll \text{U}$ this hopping term will allow an even further decrease in energy by allowing virtual processes $i \rightarrow j \rightarrow i$ that localized the electrons withing their neighbor sites. For $t \ll \text{U}$ the dynamics of the system can be described withing the low energy subspace only, by doing so we can obtain an effective spin Hamiltonian $\hat{H}_{\text{eff}} = \sum{\langle i,j \rangle} J \bs{S}_i\cdot\bs{S}_j$ where $J>0$ describing an antiferromagnetic system. A rigorous derivation of this will be offered in the next Chapter, however we can intuitively understand this result in the following way: the virtual processes $i \rightarrow j \rightarrow i$ will be allowed when the electrons at sites $i$ and $j$ have opposite spin (otherwise it would contradict the Pauli principle), therefore states where neighbor sites have opposite spins will have a lower energy due to the delocalization mechanism. Additionally, the constant $J$ will be proportional to the amplitude of the process $i \rightarrow j \rightarrow i$. We can already see that this amplitude is $2\frac{t^2}{\text{U}}$, $t^2$ due to two hopping processes, $\frac{1}{\text{U}}$ due to the energy of the intermediate state and the factor $2$ due to the two spin possibilities.

In the next Chapter and in the rest of this thesis we will consider the effects of spin orbit interaction (SOI) in strongly correlated systems. For this reason we will have an additionally spin dependent hopping term in \ref{HH} arising from the SOI. This additional term will lead to a DMI when deriving the effective spin Hamiltonian in the $t \ll \text{U}$ limit.

### 2019-03-16 Chapters/Chapter3.tex
Add more references and short summary in the intro
minor thesis changes

\chapter{Effective Spin Hamiltonian}

@@

\begin{subsection}{Hubbard model}

In the absence of spin-orbit interaction the model \ref{Hubbard} becomes the well-known Hubbard model: $\hat{H} = -\sum_{\langle i,j \rangle, \sigma} t_{ij} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + \text{U} \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$. This model has been extensively studied since its proposal in 1963 by Hubbard \cite{Hubbard1963}. Despite its simplicity the ground state has no analytic form in more than one dimensions, and different computational methods must be used to study it. Originally the Hubbard model was proposed to describe electron correlations in $d$ or $f$ bands in transition metals. The result was a simplified model to predict metal-insulator transitions. However, it exhibits much richer phenomena. It can be considered an impoved tight binding model, describing Mott insulators, which cannot be describied by band theory (i.e. they should be conductors according to band theory, but electron-electron interactions keep away from). Also, the attractive Hubbard model (where $\text{U} < 0$) can be considered an effective model relevant for understanding high temperature superconductivity \cite{Alexandrov1981}.

>>>>

\chapter{Effective Spin Hamiltonian}
\label{ChapEffectiveSpin}

@@

This Hamiltonian is identical to the Hubbard Hamiltonian \ref{HH} adding a spin dependent hopping, which arises from the SOI.
